---
title: "Descriptive Analysis and Clustering"
author: "Sirio Sobrino"
date: "25/11/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Reading data

We will work with the df_states data.set from the campus csv

```{r}
library(data.table)

df<-fread('../df_states.csv', stringsAsFactors = T)

str(df)
head(df,5)

```

### Geo-visualization of data

```{r}
library(leaflet)

map <- leaflet(df) %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lng=~center.x, lat=~center.y,
             popup=~name) 
map

```




### Missingness Analysis

```{r}
sapply(df, function(var) return(sum(is.na(var))))
```

#### Missingness Map
```{r}
library(Amelia)
missmap(df, legend = TRUE, col = c("black", "gray"), main='Missingnes map for df_states',y.labels = df$name, y.cex = 0.4)
```


#### Rebuilding the states data frame (with no NAs)
```{r}
data("state")
df<-data.table(name=rownames(state.x77),
               state.x77,
               center.x=state.center$x, 
               center.y=state.center$y)
head(df)
```


#### Visualizing Socioeconomic Indicators
```{r}
map <- leaflet(df) %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addCircleMarkers(lng=~center.x, lat=~center.y,
                   popup=~paste(name,Murder, sep=' '),
                   radius = ~Murder)
map

```

#### Visualizing Correlations among the numeric variables
```{r}
library(corrplot)

corrplot(cor(df[, sapply(df, is.numeric), with=F]))
```




#### Re-visualizing correlations, this time interactive:
```{r}
library(d3heatmap)

d3heatmap::d3heatmap(cor(df[, sapply(df, is.numeric), with=F]))

```


#### create a subset with some variables of interest for clustering
```{r}
data_cluster<-scale(df[, sapply(df, is.numeric), with=F])  # we scale the variables to mean=0, sd=1
head(data_cluster)
```


#### clustering the states based on numeric variables
```{r}
k=5
kmeans_cluster<-kmeans(data_cluster,centers = k)
print(kmeans_cluster)
```


#### Bring the cluster to the dataframe, and plot states by cluster-membership factor
```{r}
df[, cluster_k5:=as.factor(kmeans_cluster$cluster)]
head(df)


colorsmap = colors()[1:length(unique(df$cluster_k5))]
groupColors = colorFactor(palette = "Set1", domain = df$cluster_k5)
map <- leaflet(df) %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addCircleMarkers(lng=~center.x, lat=~center.y,
                   popup=~paste(name,cluster_k5, sep=' '),
                   color=~groupColors(cluster_k5),
                   group=~cluster_k5)
map


```


#### Another clustering technique: Hierarquical Clustering
```{r}
d <- dist(data_cluster, method = "euclidean")

fit <- hclust(d, method = "ward.D") 
plot(fit, labels=df$name)


rect.hclust(fit, k = k, border=1:k)

df[, cluster_h:=factor(cutree(fit, k = k))]


colorsmap = colors()[1:length(unique(df$cluster_h))]
groupColors = colorFactor(palette = "Set1", domain = df$cluster_h)
map <- leaflet(df) %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addCircleMarkers(lng=~center.x, lat=~center.y,
                   popup=~paste(name,cluster_h, sep=' '),
                   color=~groupColors(cluster_h),
                   group=~cluster_h)
map


```


#### Creating a Linear Model to explain Income:
```{r}
lm1<-lm(Income~., data=df[, !c('name','cluster_k5','cluster_h','center.x','center.y')])

print(summary(lm1))

# calculating the fitted values
df[, lm_fit:=fitted(lm1)]



df_fit<-melt(df[, .(name,Income,lm_fit)], id.vars = 'name')


library(ggplot2)
ggplot(df_fit, aes(x=name, y=value, group=variable,colour=variable))+geom_line(lwd=1, alpha=0.7)+
  scale_color_manual(values=c('black','red'))+
  theme(axis.text.x=element_text(angle=90))

```