---
title: "Descriptive Analysis"
author: "Sirio Sobrino"
date: "25/11/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Loading data

We will work with the df data.set from the carData package

```{r}
library(carData)
library(data.table)
data(Salaries)

df<-data.table(Salaries)

str(df)
head(df,5)

```

### Summarizing variables

```{r}
summary(df[, names(df)[sapply(df, is.numeric)], with=F])
summary(df[, names(df)[sapply(df, is.factor)], with=F])
```

### Correlation matrix from numeric variables

```{r}
library(corrplot)
m<-cor(df[, names(df)[sapply(df, is.numeric)], with=F]); m
corrplot(m)
```

We can also create a mixed correlation and scatterplot matrix in just one plot
```{r}
library(GGally)
ggpairs(df[, names(df)[sapply(df, is.numeric)], with=F])
```

All tree variables are positive corrrelated among them, with an almost linear 
relation between years of service and years since PhD


* Years since PhD is a better predictor for salary, with a correlation above +40% with salaries


### Does gender matter?

#### Salaries distribution by gender
```{r, echo=FALSE, fig.width=12, fig.height=6}
library(ggplot2)
theme_set(theme_minimal(base_size=16))
options(scipen = 999)
library(gridExtra)
p1<-ggplot(df, aes(y=salary,colour=sex))+geom_boxplot()
p2<-ggplot(df, aes(x=salary,colour=sex))+geom_density()

grid.arrange(p1,p2, ncol=2)

```


```{r}
print('Salary Quantiles for Males')
round(quantile(df[sex=='Male']$salary, prob=seq(0,1,0.2)))
print('Salary Quantiles for Females')
round(quantile(df[sex=='Female']$salary, probs=seq(0,1,0.2)))
```

Apparently it does..  


#### Let's perform a t-test to contrast if average salaries are statistically diffrerent depending on the proffesor's gender
```{r}
print(t.test(x=df[sex=='Male']$salary,
             y=df[sex=='Female']$salary))
```
At a 95% level of confidence, the positive difference on males' is statistically significant.  


#### But we also have to account for the rank of proffesor and the discipline
```{r, echo=FALSE, fig.width=12, fig.height=6}
p1<-ggplot(df, aes(y=salary,colour=rank))+geom_boxplot()
p2<-ggplot(df, aes(x=salary,colour=rank))+geom_density()
grid.arrange(p1,p2, ncol=2)

p1<-ggplot(df, aes(y=salary,colour=discipline))+geom_boxplot()+scale_color_manual(values = c('orange','aquamarine'))
p2<-ggplot(df, aes(x=salary,colour=discipline))+geom_density()+scale_color_manual(values = c('orange','aquamarine'))

grid.arrange(p1,p2, ncol=2)

```


#### Creating a Multiple Linear Model to explain Salaries:
```{r}
lm1<-lm(salary~yrs.since.phd+sex, data=df)

print(summary(lm1))

```

Both variables are significant at a 90% level but only yrs.since.phd at a 95% confidence.


* The estimated increase on salary per additional year of experience since PhD is about +$1000.
* The estimated difference on salaries between a male and a femal, controlling for experience since PhD, is about +$8000.

 
Though, the power of these two variables to explain the total variability on salaries is quite low, with an $R^2$ of 18%.

